---
title: "Ch. 12 Data Wrangling - Tidy Data"
author: "Jerri Schorr"
date: 10-29-2019
output: github_document
---

##### **Chapter 12**

###### **12.1**

* This chapter will teach you about the **tidy r** package
```{r include=FALSE}
library(tidyverse)
```

###### **12.2**

* There are three interrelated rules which make a dataset tidy:
    + Each variable must have its own column.
    + Each observation must have its own row.
    + Each value must have its own cell.
    
* 2 simple steps:
    + Put data in a **tibble**
    + Put variables in columns
    
* 2 main advantages
    + consistent uniform data makes it easy to use tidy tools
    + putting variables into columns means that R can treat rows of data like a vector

###### **12.3**

* Data in the real world is messy and needs tidying first
    + Sometimes the variables are not in columns but **values** are in the columns
        + You can use the **gather** function
    + Sometimes an observation is spread accross a few rows
        + You can use the **spread** function
        
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`)
```

###### **12.4**

* Sometimes we need to separate or unite values within a column
    + we can use the **seperate** or **unite** function
    
```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
```
    
###### **12.5**

* Values can be missing in 2 ways
    + Explicity (NA)
    + Implicitly (no data represented)
     
* we can use the **complete** function to fix this issue. 
    + it will take the tibble and fill blank values with NAs
    
* we can use the **fill** function to replace NAs in rows with info from the row above

###### **12.6**

* case study 

###### **12.7**

* There are two main reasons to use other data structures:
    + Alternative representations may have substantial performance or space advantages.
    + Specialised fields have evolved their own conventions for storing data that may be quite different to the conventions of tidy data.

* Either of these reasons means you’ll need something other than a tibble (or data frame).


